A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-19 manuscript.Rmd
Add citations. (#76)
* Add citations.

* fix #53

* remove faulty coded chars and abstracts

* add counter paper
#49

* place citations

Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
You may be familiar with Microsoft Words' "Track Changes"-Feature.
Version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder <!-- Aaron, reference needed-->.
This results in changes to your notes as follows:

@@

<!--I am troubled by how we should recommend to learn Git-->
Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
Other than checking technical and statistical aspects of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
With simulated data you can check if the expected data and the planned analysis match.
First of all, simulated data can be used as an initial check that the code for the planned analysis runs without error. 
This is particularly important if you want to preregister pieces of your analysis before data were collected.
For example, we can use simulated data to check whether a given data analysis outputs parameter estimates when we simulate missing data values.
Second, Monte Carlo simulations allow us to form realistic expectations about the unbiasedness and precision of parameter estimates and it can help with sample size planning to obtain a sensible study design. 
For a variety of standard models, there are analytic strategies to evaluate precision or statistical power, however, Monte Carlo simulation offers more flexibility in terms of mode realistic modeling conditions (e.g., missing data or non-normality).
For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data. 
No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
A simulation forces us to deal with these assumptions and their consequences explicitly.
If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one-fits-all solution.
Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.

If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Random Number Generation"](#prng).
To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.

@@

As a starting point, we can use already published information.
@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
They report a gender difference (at the manifest level), which depending on the study ranges between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

@@

Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
The following code simulates data from the population model `model_truth` with 1000 cases in each group.

@@

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption has also been incorporated into our simulation of the data.
There is a good chance that this assumption is violated.
We therefore adjust our analysis plan to include tests for measurement invariance.
We propose the following procedure. 
After fitting the model with strong measurement invariance <!--AP: citation needed-->, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our modeling assumptions, it is also unlikely that the data are perfectly normally distributed.
First, we have rating scale items and thus our observations are categorical.
Second, @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity of our illustration, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).

@@

Note, that this certainly results in different numbers each time we run it (see ["Random Number Generation"](#prng)).
Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.

@@

### Random Number Generation (optional) {#prng}

Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`. Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The problem is that concurrent random number streams are correlated, because they repeat after some time.
Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:

@@

The second advantage is transparency.
A preregistration offers the interested reader more insight into the research process than a paper alone could.
Some [citation!] argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
While such an argument has some merits, we believe that often transparency outweighs these concerns.
Deviation from a preregistration is still possible.
If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study, and any deviation will simply be explicitly marked as such.
A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
PAC essentially reverses the process of empirical research.
Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

@@

<!--Hands on: Introduce R Markdown-->
As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."

Above, we have already suggested [Markdown] as a universal and portable text format. 
R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).

To create an R Markdown document, click on:

@@

You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

@@

<!-- Hands-on: check changed hash -->
A prerequisite to reproducibility is the availability of code and data.
We use containers to guarantee that all code components are available and work as expected.
Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
You may find yourself in a situation where you can make the data only available on request.
A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
If the data changes, so does its fingerprint or hash.
The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.

>>>>

Another advantage of Markdown is that because of its alikeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
You may be familiar with Microsoft Words' "Track Changes"-Feature.
Version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder [@flake2020].
This results in changes to your notes as follows:

@@

<!--I am troubled by how we should recommend to learn Git-->
Git can be challenging to learn, because it requires a rather complex mental model.
However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to utilize Git for simple version control.
There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
One has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.

If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.

@@

As a starting point, we can use already published information.
@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
They report a gender difference (at the manifest level) depending on the study between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

@@

Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], faux [@faux], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
The following code simulates data from the population model `model_truth` with 1000 cases in each group.

@@

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption is also baked into our simulation of the data.
There is a good chance that this assumption is violated.
We, therefore, adapt our analysis to plan for this contingency.
After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
Anyhow, remember that what we consider realistic for our case could be unfit for your case.

@@

Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.

@@

### Randome Number Generation (optional) {#prng}

Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The problem is that concurrent random number streams are correlated, because they repeat after some time.
Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:

@@

The second advantage is transparency.
A preregistration offers the interested reader more insight into the research process than a paper alone could.
@szollosi_is_2020 argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
While such an argument has some merits, we believe that often transparency outweighs these concerns [@nosekPreregistrationHardWorthwhile2019].
Deviation from a preregistration is still possible.
If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study, and any deviation will simply be explicitly marked as such.
A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code [@sep-ambiguity].
Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
PAC essentially reverses the process of empirical research.
Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

@@

<!--Hands on: Introduce R Markdown-->
As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
We already learned about [Markdown], to write text.
R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
To create an R Markdown document, click on:

@@

You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

@@

<!-- Hands-on: check changed hash -->
A prerequisite to reproducibility is the availability of code and data [@akhlaghi2020towards].
We use containers to guarantee that all code components are available and work as expected.
Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs] and well documented [@codebook], but this is not always feasible.
Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
You may find yourself in a situation where you can make the data only available on request [e.g. @share].
A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
If the data changes, so does its fingerprint or hash.
The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.
